---
title: "Lab1"
author: "Nathaniel Gunter"
date: "1/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```
# Lab 1
## Task 1

Downloaded "Dataxls" and unzipped into LAB1.
Then obtain the working directory (to check) using:
```{r checkDirectory}
getwd()
```

## Task 2

Saved DDT.xls as DDT.csv using Microsoft Excel. 
Read the data using:
```{r readTable}
ddt = read.csv("DDT.csv")
```
Then check the first six lines using:
```{r head}
head(ddt)
```

## Task 3 

To better understand charactaristics of the data set, we use:
```{r summarize}
knitr::kable(summary(ddt))
```

From this summary, we see the following properties.
The qualitative variables in ddt are RIVER and SPECIES. 
The quantitative variables are MILE, LENGTH, WEIGHT, and DDT.
There are three different species in this data set. 

Now we subset the data for SPECIES==LMBASS and WEIGHT larger than 800 grams:
```{r LMBASSWeight}
lmbass800 = with(ddt, ddt[SPECIES == 'LMBASS' & WEIGHT > 800,])
head(lmbass800)
```

Then subset the data for RIVER==SCM and DDT>4.0:
```{r SCMDDT}
scmddt = with(ddt, ddt[RIVER=='SCM' & DDT>4.0,])
head(scmddt)
```

CLICKER QUESTIONS:  
Q.1: What is the mean length of fish in the ddt data frame?  
  A.1: The mean length of fish in ddt is 45.00 cm.  
Q.2: What is the standard deviation of the weight of fish in the ddt data frame?  
  A.2: The standard deviation of weight in ddt is:
```{r stnd}
sd(ddt$WEIGHT,na.rm = FALSE)
```  
Q.3: Is this the plot of LENGTH vs WEIGHT from the ddt data frame?  
  A.3: No, the plot has its axes flipped compared to the example.
```{r length v weight}
ggplot(ddt, aes(x=WEIGHT, y=LENGTH))+geom_point()+ggtitle("LENGTH vs WEIGHT from data frame ddt")
```  
Q.4: If v = 1:20 what is the last value of v/20?  
  A.4: 1.00  
  
## Task 4 

Make a table of the rivers.  
```{r riverTable}
table(ddt$RIVER)
```
Make a color coded barplot of the rivers.  
```{r riverPlot}
ddt %>% ggplot(aes(x=RIVER, fill=RIVER))+geom_bar()+ggtitle("Nathaniel Gunter")
```  

Make a table of the rivers crossed with the species of fish.  

Make a barplot of the rivers crossed with the species of fish.  

## Task 5 

Make a pie chart of Fish SPECIES.  
```{r speciesPie}
speciesPie<- ddt%>%ggplot(aes(x="",y=SPECIES, fill = SPECIES), )+theme(axis.text.y = element_blank())+geom_bar(width=1, stat="identity")+ggtitle("Nathaniel Gunter") + theme(axis.text = element_blank(), axis.ticks=element_blank(), axis.title.x = element_blank(), axis.title.y=element_blank(), axis.line=element_blank())+coord_polar("y",start=0)
speciesPie
```  

Make a pie chart of RIVERS.  

```{r riversPie}
riversPie <- ddt%>%ggplot(aes(x="",y=RIVER, fill=RIVER), )+geom_bar(width=1, stat="identity")+ggtitle("Nathaniel Gunter")+ theme(axis.text = element_blank(), axis.ticks=element_blank(), axis.title.x = element_blank(), axis.title.y=element_blank(),axis.line=element_blank())+coord_polar("y",start=0)
riversPie
```  


## Task 6  

Make a boxplot of the variable DDT.  

```{r ddtbox}
ddtbox <- ddt%>%ggplot(aes(x = "",y=DDT),)+geom_boxplot()+ggtitle("Nathaniel Gunter")
ddtbox
```  

Make a boxplot of the variable WEIGHT.  

```{r weightbox}
weightbox <- ddt%>%ggplot(aes(x = "", y = WEIGHT),)+geom_boxplot()+ggtitle("Nathaniel Gunter")
weightbox
```  

Make a boxplot of the variable LENGTH.  

```{r lenbox}
lenbox <- ddt%>%ggplot(aes(x="",y=LENGTH),)+geom_boxplot()+ggtitle("Nathaniel Gunter")
lenbox
```  

## Task 7  

Make a coplot of LENGTH vs. WEIGHT given RIVER.  

```{r lenweightcoplot}
coplot(LENGTH ~ WEIGHT | RIVER, data = ddt)
```  

Make a coplot of DDT vs. WEIGHT given SPECIES.  

```{r ddtweightcoplot}
coplot(DDT ~ WEIGHT | SPECIES, data = ddt)
```  

## Task 8  

Make the following boxplots using 'ggplot'.  

```{r boxplots}
boxplots <- ddt%>%ggplot(aes(x=SPECIES, y=WEIGHT, fill=RIVER))+geom_boxplot()+ggtitle("Nathaniel Gunter")
boxplots
```  

Make the following violin plots.  

```{r violinplots}
violinplots <-ddt%>%ggplot(aes(x=RIVER,y=LENGTH,fill=SPECIES))+geom_violin()+ggtitle("Nathaniel Gunter")
violinplots
```  

Make the following scatter plots.  

```{r scatterplots}
scatterplots<-ddt%>%ggplot(aes(x=WEIGHT,y=LENGTH,colour=SPECIES))+geom_point()+ggtitle("Nathaniel Gunter")
scatterplots
```  
